---
title: "Setup"
output: rmarkdown::html_vignette
description: |
  Getting R and tameDP installed
vignette: >
  %\VignetteIndexEntry{Setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

This vignette provides an overview for getting started working with `tameDP`. The first time you setup R or new packages, you can run into some hurdles, but one you have things setup, everything should be smooth sailing.

### Installing R

In order to use `tameDP`, you will need to install R, an open-source statistics package. You can do this on your USAID GFE, by going to Start > Microsoft Endpoint manager > Software Center. Once there, you can select the Application called "R for Windows" and click "Install". If working on a personal machine, you can install R from [CRAN](https://cran.r-project.org/). Select "Download R for Windows" and then "base" and follow the instructions for installing that pop up when you launch the .exe file from your downloads.


### Installing RStudio

R is the underlying software we're using, but the user interface leaves much to be desired. RStudio used to be available in Software Center as well, but M/CIO removed around October 2021. **You don't need RStudio to run `tameDP`**, but if desired, you can submit a Help Desk ticket with M/CIO and they will install it for you. If working on a personal machine, you can install RStudio from [this site](https://www.rstudio.com/products/rstudio/download/). Select "Download" for the free version and follow the instructions for installing that pop up when you launch the .exe file from your downloads.

### Installing Required Packages

With proprietary programs like Excel, Tableau, or Stata, all of your functions are built into the software. With open-source packages on the other hand, we need to load different libraries or packages that are written by other organization or individuals users. We wrote the package `tameDP` and it relies on the work others have done as well, so we install those packages first to make things easy. To install a package from CRAN, we use the command `install.packages()`. Copy and paste the code chunk below to install all the required packages.

```{r install packages, eval=F, echo=T}
install.packages("tidyverse")
install.packages("remotes")
install.packages("curl")
install.packages("getPass")
```

### Installing tameDP

Our package is hosted on a site called GitHub rather than CRAN. To install our package now that you have the other required packaged installed, you will use a function from the `remotes` package.

```{r install tameDP, eval=F, echo=T}
remotes::install_github("USAID-OHA-SI/tameDP", build_vignettes = TRUE)
```

With all the software and packages installed, you're ready to start using `tameDP`.

